---
title: "Homework 4"
format: pdf
editor: visual
---
## Task 1: Conceptual Questions

### Question 1: What is the purpose of the lapply() function? What is the equivalent purrr function?

The purpose of the lapply() function is to apply a function to each element in a list in R. The equivalent purrr function is map().

### Question 2: Suppose we have a list called my_list. Each element of the list is a numeric data frame (all columns are numeric). We want use lapply() to run the code cor(numeric_matrix, method = "kendall") on each element of the list. Write code to do this below! (I’m really trying to ask you how you specify method = "kendall" when calling lapply())

```{r}
lapply(my_list, cor, method = "kendall")
```

### Question 3: What are two advantages of using purrr functions instead of the BaseR apply family?

1. Greater consistency between functions when using the purrr package

2. Almost every purrr function is type stable, meaning you are easily able
to predict the type of data output you will receive from the function name. 

### Question 4: What is a side-effect function?

A side-effect function does not actually change the data, instead the goal 
is to just produce something. Some examples include; plot(), print(), and write.csv().

### Question 5: Why can you name a variable sd in a function and not cause any issues with the sd function?

Variable names within functions create a temperorary environment that does not overwrite other functions. 

## Task 2: Writing R Functions
### Question 1: Write a basic function (call it getRMSE()) that takes in a vector of responses and a vector of predictions and outputs the RMSE.
```{r}

getRMSE <- function(vector_resp, vector_pred){
  
  diff_sq <- (vector_resp - vector_pred)^2
  
  MSE <- mean(diff_sq, na.rm = TRUE)
  
  RMSE <- sqrt(MSE)
  
  return(RMSE)
  
}
```

### Question 2: Testing getRMSE()
```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

# No NA

getRMSE(resp, pred)

# With NA

resp_missing <- resp
resp_missing[c(13, 20)] <- NA_real_

getRMSE(resp_missing, pred)

```

### Question 3: Write a function called getMAE() that follows the specifications of the getRMSE() function.

```{r}

getMAE <- function(vector_resp, vector_pred){
  
  abs_diff <- abs(vector_resp - vector_pred)
  
  MAE <- mean(abs_diff, na.rm = TRUE)
  
  return(MAE)
}

```

### Question 4: Testing getMAE()
```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

#No NA

getMAE(resp, pred)

# With NA

resp_missing <- resp
resp_missing[c(13, 20)] <- NA_real_

getMAE(resp_missing, pred)

```
### Question 5: Create a wrapper function 
```{r}
wrapper <- function(vector_resp, vector_pred, metrics = c("RMSE", "MAE"))
                    {
  
  if((!is.vector(vector_resp)) | (!is.atomic(vector_resp))| ((!is.numeric(vector_resp))) | (!is.vector(vector_pred)) | (!is.atomic(vector_pred))| (!is.numeric(vector_pred))){
    return("Both inputs must be numeric atomic vectors.")
  } else{
  
  result <- list()
  
  if("RMSE" %in% metrics){
    result$RMSE <- getRMSE(vector_resp, vector_pred)}
  
  if("MAE" %in% metrics){
    result$MAE <- getMAE(vector_resp, vector_pred)}
  
 return(result)}
  
}
  
```

### Question 6: Test wrapper function
```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

# No NA (Individual)

wrapper(resp, pred, metric = "RMSE")

wrapper(resp, pred, metric = "MAE")

# No NA (Both Metrics)

wrapper(resp, pred)

# With NA

resp_missing <- resp
resp_missing[c(13, 20)] <- NA_real_

wrapper(resp_missing, pred, metric = "RMSE")

wrapper(resp_missing, pred, metric = "MAE")

wrapper(resp_missing, pred)

# Incorrect data

resp_incorrect <- data.frame(1:100)

wrapper(resp_incorrect, pred)
```

